The Small World Network Effect in Open Source Project Teams \author{Kevin Peterson\\ \small \texttt{pete1968@umn.edu} }

### 2013-11-18 paper/main.tex
Fixed bad reference links.

\begin{figure}
\label{fig:example_bipartite_graph}
\centering
\begin{tikzpicture}
\tikzset{P/.style={circle,fill=lightgray!20}}
\tikzset{C/.style={circle,fill=black!20}}

@@

  \draw ({p1}) -- ({c1});
  \draw ({p1}) -- ({c2});
  \draw ({p2}) -- ({c2});;
  \draw ({p3}) -- ({c2});
  \draw ({p3}) -- ({c3});
  \draw ({p4}) -- ({c3});
  \draw ({p4}) -- ({c4});
  \node[draw=none,fill=none,rectangle,left=1mm of p1] {Projects}; 
  \node[draw=none,fill=none,rectangle,right=1mm of c1] {Contributors}; 
\end{tikzpicture}
\caption{An example bipartite graph of Projects and Contributors}
\end{figure}

@@

\begin{figure}
\label{fig:example_bipartite_projection_graph}
\centering
\begin{tikzpicture}
\tikzset{P/.style={circle,fill=lightgray!20}}
\tikzset{C/.style={circle,fill=black!20}}

@@

  \draw ({p1}) -- ({p3});
  \draw ({p1}) -- ({p2});
  \draw ({p1}) -- ({p2});
  \draw ({p2}) -- ({p3});
  \draw ({p3}) -- ({p4});
\end{tikzpicture}
\caption{A $P$ projection of the figure \ref{fig:example_bipartite_graph} bipartite graph}
\end{figure}

@@

\begin{figure}
\label{fig:example_graphs}
\centering
\begin{tikzpicture}
  [scale=.5,auto=left,every node/.style={circle,fill=gray!20}]
  \node (n1) at (15,8) {1};
  \node (n2) at (12,6)  {2};
  \node (n3) at (9,5)  {3};
  \node (n4) at (6,5)  {4};
  \node[draw=none,fill=none,rectangle,right=1mm of n1] {$C^\Delta = 0.0$};

@@

  \draw ({n1}) -- ({n2});
  \draw ({n3}) -- ({n4});
  \draw ({n1}) -- ({n4});
  \draw ({n2}) -- ({n4});
\end{tikzpicture}
\caption{A caption.}
\end{figure}

\textit{Small-World-ness}\\

\[\gamma^{\Delta}_g = \frac{C^{\Delta}_g }{C^{\Delta}_{rand} } \]

\[\lambda_g = \frac{L_g }{L_{rand} } \]

\[S^\Delta = \frac{\gamma^{\Delta}_g }{\lambda_g} \]

@@

To validate the hypothesis, we expect to find that medium values of $C^\Delta$ will be correlated with more popular projects. To explore this possibility, we first compare the popularity given $C^\Delta$ values as shown in figure \ref{fig:cc_popular}. The $C^\Delta$ values do display the U-shaped characteristic as compared to popularity, which aligns with the findings of Uzzi \textit{et. al}\cite{uzzi2005collaboration}. It is important to note, however that, the standard deviation is high, as shown in both \ref{fig:summary_stats} and the significant difference between the mean and median results.

\begin{figure}
\label{fig:cc_popular}
\begin{center}
\caption{$C^\Delta$ of Popular Projects}
\includegraphics[width=0.5\textwidth]{images/freecode-popular.png}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{$C^\Delta$ of Unpopular Projects}
\includegraphics[width=0.5\textwidth]{images/freecode-unpopular.png}
\end{center}
\end{figure}

\begin{figure}
\label{fig:q_distribution}
\begin{center}
\caption{$Q$ Distribution}
\includegraphics[width=0.5\textwidth]{images/freecode-q-histo.png}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{$S^\Delta$ Distribution}
\includegraphics[width=0.5\textwidth]{images/freecode-smallworld-histo.png}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{$C^\Delta$ and Project Popularity}
\includegraphics[width=0.5\textwidth]{images/freecode-graph.png}
\end{center}
\end{figure}

@@

>>>>

\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{P/.style={circle,fill=lightgray!20}}
\tikzset{C/.style={circle,fill=black!20}}

@@

  \draw ({p1}) -- ({c1});
  \draw ({p1}) -- ({c2});
  \draw ({p2}) -- ({c2});;
  \draw ({p3}) -- ({c2});
  \draw ({p3}) -- ({c3});
  \draw ({p4}) -- ({c3});
  \draw ({p4}) -- ({c4});
  \node[draw=none,fill=none,rectangle,left=1mm of p1] {Projects}; 
  \node[draw=none,fill=none,rectangle,right=1mm of c1] {Contributors}; 
\end{tikzpicture}
\caption{An example bipartite graph of Projects and Contributors}
\label{fig:example_bipartite_graph}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{P/.style={circle,fill=lightgray!20}}
\tikzset{C/.style={circle,fill=black!20}}

@@

  \draw ({p1}) -- ({p3});
  \draw ({p1}) -- ({p2});
  \draw ({p1}) -- ({p2});
  \draw ({p2}) -- ({p3});
  \draw ({p3}) -- ({p4});
\end{tikzpicture}
\caption{A $P$ projection of the figure \ref{fig:example_bipartite_graph} bipartite graph}
\label{fig:example_bipartite_projection_graph}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture}
  [scale=.5,auto=left,every node/.style={circle,fill=gray!20}]
  \node (n1) at (15,8) {1};
  \node (n2) at (12,6)  {2};
  \node (n3) at (9,5)  {3};
  \node (n4) at (6,5)  {4};
  \node[draw=none,fill=none,rectangle,right=1mm of n1] {$C^\Delta = 0.0$};

@@

  \draw ({n1}) -- ({n2});
  \draw ({n3}) -- ({n4});
  \draw ({n1}) -- ({n4});
  \draw ({n2}) -- ({n4});
\end{tikzpicture}
\caption{A caption.}
\label{fig:example_graphs}
\end{figure}

\textit{Small-World-ness}\\
The idea of a \textit{small world} can be formalized by the following equations\cite{humphries2008network}.

First, let $\gamma^{\Delta}_g$ equal the ratio of the clustering coefficient of a given graph and a random graph.
\[\gamma^{\Delta}_g = \frac{C^{\Delta}_g }{C^{\Delta}_{rand} } \]

Next apply a similar pattern to the average path length $\lambda_g$.
\[\lambda_g = \frac{L_g }{L_{rand} } \]

Finally, the ratio of the above calculations yields the small-world-ness, or $S^\Delta$, of the graph
\[S^\Delta = \frac{\gamma^{\Delta}_g }{\lambda_g} \]

We assume for throughout that any graph exhibiting $S^\Delta > 1$ is a small-world graph.

\textit{Successful vs Unsuccessful}\\
To further explore the data, a comparion was made between \textit{successful} and \textit{unsuccessful} projects was made. We define \textit{successful} as any project with {$PopularityScore > 100$}, and \textit{unsuccessful} as any project with {$PopularityScore < 50$}. Given this view of the data, we can then make assertions about how we expect these types of projects to behave, given the stated hypothesis. More specifically, we can examine not only if the u-shaped curve holds for \textit{popular} projects, but if the inverse holds for \textit{unsuccessful} ones.

@@

To validate the hypothesis, we expect to find that medium values of $C^\Delta$ will be correlated with more popular projects. To explore this possibility, we first compare the popularity given $C^\Delta$ values as shown in figure \ref{fig:cc_graph}. The $C^\Delta$ values do display the U-shaped characteristic as compared to popularity, which aligns with the findings of Uzzi \textit{et. al}\cite{uzzi2005collaboration}. It is important to note, however that, the standard deviation is high, as shown in both \ref{fig:summary_stats} and the significant difference between the mean and median results.

\begin{figure}
\begin{center}
\caption{$C^\Delta$ and Project Popularity}
\includegraphics[width=0.5\textwidth]{images/freecode-graph.png}
\end{center}
\caption{A caption.}
\label{fig:cc_graph}
\end{figure}

Popular projects do appear to conform to the results found by Uzzi\cite{uzzi2005collaboration}. As show in figure \ref{fig:cc_unpopular}, $C^\Delta$ near the extremes (0.0 or 1.0) show a relatively low likelihood of being popular. As $C^\Delta$ values approach mid-range values, the probability of a project being a success increases greatly.

\begin{figure}
\begin{center}
\caption{$C^\Delta$ of Popular Projects}
\includegraphics[width=0.5\textwidth]{images/freecode-popular.png}
\end{center}
\caption{A caption.}
\label{fig:cc_popular}
\end{figure}

The inverse is true of unpopular projects. As $C^\Delta$ near the extremes (0.0 or 1.0), a high probability of being unsuccessful is observed. Mid-valued $C^\Delta$ values indicate a substantially lower probability of being unsuccessful.

\begin{figure}
\begin{center}
\caption{$C^\Delta$ of Unpopular Projects}
\includegraphics[width=0.5\textwidth]{images/freecode-unpopular.png}
\end{center}
\caption{A caption.}
\label{fig:cc_unpopular}
\end{figure}

\begin{figure}
\begin{center}
\caption{$Q$ Distribution}
\includegraphics[width=0.5\textwidth]{images/freecode-q-histo.png}
\end{center}
\caption{A caption.}
\label{fig:q_distribution}
\end{figure}

\begin{figure}
\begin{center}
\caption{$S^\Delta$ Distribution}
\includegraphics[width=0.5\textwidth]{images/freecode-smallworld-histo.png}
\end{center}
\caption{A caption.}
\label{fig:sw_distribution}
\end{figure}

@@

A futher area of study would be how this relates to previously defined \textit{success factors} in projects\cite{cooke2002real}.
